---
title: "PrimarySchool"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PrimarySchool}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## The network of contacts in a primary school

The network of contacts in a primary school was collected an first analyzed in "High-Resolution Measurements of Face-to-Face Contact Patterns in a Primary School" by J. Stehlé et al. in PLOS ONE (2011). This data set records physical interactions between $226$ children and $10$ teachers from the same primary school over the course of a day. The network data was collected using a system of sensors worn by the participants. This system records the duration of interactions between two individuals facing each other at a maximum distance of one and a half meters.
```{r load data}
library(Matrix)
library(gsbm)
library(igraph)
library(RColorBrewer)

data(PrimarySchool)
summary(PrimarySchool)
```
